#Personal Movement Data Analysis
###Assignment: Course Project 1
______________

##Loading and preprocessing the data

Loading the data  
```{r cashe=TRUE} 
Data<-read.csv(file = "activity.csv")
```

______________
##What is mean total number of steps taken per day?
 
1. Calculating the total number of steps taken per day

```{r message=FALSE}
library(dplyr)
grouped_by_day<-group_by(Data, date)
steps_per_day<-summarize(grouped_by_day, steps_per_day=sum(steps))
```

2. Making the table presenting the total number of steps taken per day
```{r results="asis"}
library(xtable)
xt1<-xtable(steps_per_day)
print(xt1, type = "html")
```

3. Making the histogram of the number of days with the given number of steps per day
```{r message=FALSE, warning=FALSE}
library(ggplot2)
plot1<-ggplot(steps_per_day, aes(steps_per_day))+geom_histogram()+labs(title="Number of days in which the given number of steps was taken")+labs(x="Number of steps", y="Count of days")
print(plot1)
```

4. Calculating and reporting the mean and median of the total number of steps taken per day
```{r}
Mean<-mean(steps_per_day$steps_per_day, na.rm = TRUE)
Median<-median(steps_per_day$steps_per_day, na.rm = TRUE)
```
The mean of the total number of stepts per day is **`r Mean`** and the median is **`r Median`**.


##What is the average daily activity pattern?

1. Calculating avarage number of steps per interval (averaged accross all days)
```{r}
grouped_by_interval<-group_by(Data, interval)
steps_per_interval<-summarize(grouped_by_interval, mean_steps_per_interval=mean(steps, na.rm = TRUE))
```

2. Making a time series plot

```{r}
plot2<-with(steps_per_interval, plot.ts(interval, mean_steps_per_interval, 
                                 type="l", 
                                 col="blue",
                                 main = "Average number of steps (accross the days)",
                                 xlab="Time interval (minutes)",
                                 ylab="Average number of steps"))
print(plot2)
```

3. Indicating 5-minute interval with maximum nuber of steps (average accross all the days)

```{r} 
interval_max<-which(steps_per_interval$mean_steps_per_interval==max(steps_per_interval$mean_steps_per_interval))
interval_max<-steps_per_interval[interval_max, 1]
```
Interval with maximum average (accross all days) number of steps is **`r interval_max`**.

##Imputing missing values

1. Counting NAs
```{r}
NA_count<-sum(is.na(Data))
```
The number of missing values in the dataset is **`r NA_count`**.

2. Filling in missing values with the mean of steps (accross all the days) of 5-minute intervals
```{r}
impData<-Data
for (i in (1:nrow(impData))) {
        if (is.na(impData$steps[i])){
                impData$steps[i]<-steps_per_interval[(which(steps_per_interval$interval==impData$interval[i])),]$mean_steps_per_interval
        }
}
```

3. Chacking if all missing values are filled in
```{r}
number_of_NAs<-sum(is.na(impData))
```
Now the number of missing data in dataset is **`r number_of_NAs`**.


4. Calculating the total number of steps taken per day in the dataset without missing data

```{r message=FALSE}
imp_grouped_by_day<-group_by(impData, date)
imp_steps_per_day<-summarize(imp_grouped_by_day, steps_per_day=sum(steps))
```
5. Making a histogram of the total number of steps taken each day
```{r message=FALSE, warning=FALSE}
plot3<-ggplot(imp_steps_per_day, aes(steps_per_day))+geom_histogram()+labs(title="Number of days in which the given number of steps was taken (no NAs it the dataset)")+labs(x="Number of steps", y="Count of days")
print(plot3)
```

6. Calculating and reporting the mean and median of the total number of steps taken per day in the dataset without missing values
```{r}
impMean<-mean(imp_steps_per_day$steps_per_day, na.rm = TRUE)
impMedian<-median(imp_steps_per_day$steps_per_day, na.rm = TRUE)
```
The mean of the total number of stepts per day in the dataset without missing values is **`r impMean`** and the median is **`r impMedian`**. That means that filling in the missing values had **almost no impact** on the mean and the median, i.e. mean changed by **`r impMean-Mean`** and median by **`r impMedian-Median`**.

##Are there differences in activity patterns between weekdays and weekends?

1. Creating a new factor variable indicating whether a given date is a weekday or weekend day.
```{r}
impData$day<-factor(weekdays(as.Date(impData$date)))
levels(impData$day)<-list("weekday"=c("poniedziałek", "wtorek", "środa", "czwartek", "piątek"),"weekend"=c("sobota", "niedziela"))
```

2. Calculating the mean of steps per interval for weekdays and weekned days (dataset with NAs filled in)
```{r}
impData_gruped_by_interval_and_day<-group_by(impData, interval, day)
mean_steps<-summarize(impData_gruped_by_interval_and_day, mean=mean(steps))
```

3. Making the time series plot with panels for weekdays and weekend days
```{r}
library(lattice)
plot4<-xyplot(mean~interval| day, data=mean_steps, 
       type="l", 
       layout=c(1,2), 
       xlab="Interval", 
       ylab="Number of steps")
print(plot4)
```
